Condensed Matter Field Theory notes

### 2017-03-09 Condensed Matter Field Theory note/CMF-Atland.tex
[minor] Some notes about Spin path integral, boson coherent state, in Condensed matter field theory.

The last problem is that, the action in the new coordinate is:
\begin{equation}
    \int_0^{r_0}\dd{r} \int_0^{2\pi}\dd{\theta} \left[
        m(\partial_r\phi)^2+V(\phi)-f\phi\right] r
\end{equation}
where I have ignored several constants, and have used the formula:
\begin{equation}
    \dd{\tau}\dd{x} = r\dd{r}\dd{\theta}
\end{equation}
For the moment ignores $-f\phi r$, the first two terms:
\begin{equation}
    \int_0^{2\pi}\dd{\theta} \int_0^{r_0}\dd{r} \left[
        m(\partial_r\phi)^2 + V(\phi)
    \right] r
\end{equation}
is still one step away from producing a $S_{\text{inst}}$ term:
\begin{equation}
    S_{\text{inst}} := \int_0^T \dd{\tau} 
    \left(\frac{m}{2}(\partial_\tau\phi)^2 + V(\phi)\right)
\end{equation}
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}

>>>>

The last problem is that, the action in the new coordinate is:
\begin{equation}
    \int_0^{r_0}\dd{r} \int_0^{2\pi}\dd{\theta} \left[
        m(\partial_r\phi)^2+V(\phi)-f\phi\right] r
\end{equation}
where I have ignored several constants, and have used the formula:
\begin{equation}
    \dd{\tau}\dd{x} = r\dd{r}\dd{\theta}
\end{equation}
For the moment ignores $-f\phi r$, the first two terms:
\begin{equation}
    \int_0^{2\pi}\dd{\theta} \int_0^{r_0}\dd{r} \left[
        m(\partial_r\phi)^2 + V(\phi)
    \right] r
\end{equation}
is still one step away from producing a $S_{\text{inst}}$ term:
\begin{equation}
    S_{\text{inst}} := \int_0^T \dd{\tau} 
    \left(\frac{m}{2}(\partial_\tau\phi)^2 + V(\phi)\right)
\end{equation}
\section{Eq. 3.50}

This equation is actually a lemma in combinatorics, see

\href{https://en.wikipedia.org/wiki/Baker%E2%80%93Campbell%E2%80%93Hausdorff_formula#An_important_lemma}{Wikipedia
Baker–Campbell–Hausdorff formula}.

It says:
\begin{align}
    \mathrm{Ad}_{e^X}Y &\equiv e^X Y e^{-X} = e^{\mathrm{ad}_X}Y  \nonumber\\
    &\equiv Y + [X,Y] + \frac{1}{2!}[X,[X,Y]] + \frac{1}{3!}[X,[X,[X,Y]]] + \cdots
\end{align}
or just
\begin{equation}
    \mathrm{Ad}_{e^X} = e^{\mathrm{ad}_X}
\end{equation}

Here the author uses the notation $[X,\,] Y \equiv \mathrm{ad}_X Y \equiv
[X,Y]$.
\section{p.138 Exercise}
Using eq.3.50 on p.138, it is straightforward to obtain:
\begin{align*}
    n_1 &= \cos{\phi}\cos{\theta}\braket{S_1} +
    \cos{\phi}\sin{\theta}\braket{S_3} -\sin{\phi}\braket{S_2} \\
    n_2 &= \cos{\phi}\braket{S_2} + \sin{\phi}\cos{\theta}\braket{S_1} +
    \sin{\theta}\sin{\phi}\braket{S_3} \\
    n_3 &= \cos{\theta}\braket{S_3} + \sin{\theta}\braket{S_1}
\end{align*}
And we have
\begin{align*}
    \braket{S_1}&=\braket{\uparrow| S_1 |\uparrow} = 0 \\
    \braket{S_2}&=\braket{\uparrow| S_2 |\uparrow} = 0 \\
    \braket{S_3}&=\braket{\uparrow| S_3 |\uparrow} = S
\end{align*}
by the representation thoery of $\mathrm{SU}(2)$.
\section{Eq. 3.52}

\begin{align}
    \braket{\partial_\tau g|g} 
    &= \braket{\uparrow|
        \partial_\tau e^{i\phi\hat{S}_3+i\theta\hat{S}_2}
        e^{-i\phi\hat{S}_3-i\theta\hat{S}_2}
        |\uparrow}
    \nonumber\\
    &= \braket{\uparrow|
        e^{i\phi\hat{S}_3+i\theta\hat{S}_2}
        i (\hat{S}_3\partial_\tau\phi + \hat{S}_2\partial_\tau\theta)
        e^{-i\phi\hat{S}_3-i\theta\hat{S}_2}
        |\uparrow }
    \nonumber\\
    &= i\left(\partial_\tau\phi\braket{g|\hat{S}_3|g} +
        \partial_\tau\theta\braket{g|\hat{S}_2|g} \right)
    \nonumber\\
    &= i(n_3\partial_\tau\phi + n_2 \partial_\tau\theta)
    \nonumber\\
    &= i S(\cos{\theta}\partial_\tau\phi + \sin{\theta}\sin{\phi}\partial_\tau\theta)
\end{align}
Then how is $\partial_\tau\theta$ missing in (3.52).

\section{Boson coherent state (p.159)}

The aim is to show:
\begin{equation}
    a \exp(\phi a^\dagger) \ket{0} = \phi\exp(\phi a^\dagger)\ket{0}
\end{equation}

Use $aa^\dagger = 1+ a^\dagger a$. And observe:
\begin{align*}
    a(a^\dagger)^n &= (a^\dagger)^{n-1}+ a^\dagger a (a^\dagger)^{n-1} \\
    &= (a^\dagger)^{n-1}+ (a^\dagger)^{n-1}+ (a^\dagger)^2 a (a^\dagger)^{n-2}
    \\
    &= 3(a^\dagger)^{n-1} + (a^\dagger)^3 a (a^\dagger)^{n-3} \\
    &\cdots \\
    &= n (a^\dagger)^{n-1} + (a^\dagger)^{n} a
\end{align*}
Then (notice $a\ket{0}=0$):
\begin{align*}
    \sum_n a \frac{\phi^n (a^\dagger)^n}{n!}\ket{0} 
    &= \sum_n \phi\, \phi^{n-1} \frac{n(a^\dagger)^{n-1}+(a^\dagger)^n a}{n!}\ket{0}\\
    &= \sum_n \phi\, \phi^{n-1} \frac{n(a^\dagger)^{n-1}}{n!}\ket{0} \\
    &= \phi\sum_n \phi^{n} \frac{(a^\dagger)^{n}}{n!}\ket{0} \\
    &= \phi\exp{\phi a^\dagger}\ket{0}
\end{align*}
\section{License}
The entire content of this work (including the source code
for TeX files and the generated PDF documents) by 
Hongxiang Chen (nicknamed we.taper, or just Taper) is
licensed under a 
\href{http://creativecommons.org/licenses/by-nc-sa/4.0/}{Creative 
Commons Attribution-NonCommercial-ShareAlike 4.0 International 
License}. Permissions beyond the scope of this 
license may be available at \url{mailto:we.taper[at]gmail[dot]com}.
\end{document}
